The Rockerverse: Packages and Applications for Containerisation with R

### 2020-06-09 rockerverse.Rmd
fix Celeste's affiliation
Added pointers to 1-Click app and webhooks

First, different packages use containers for the **deployment of R and \pkg{Shiny} apps**.
\pkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications \citep{cran_shiny}.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open-source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarios with multiple apps.
ShinyProxy itself is written in Java to accommodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy can be achieved, e.g., via [GitLab pipelines](https://docs.gitlab.com/ee/ci/pipelines.html).
A pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed to GitLab's container registry.
The pipeline finishes with triggering a webhook on the server, where the
deployment script is executed.
The script updates configurations, pulls the new Docker images, and restarts the services.

@@

The ordering of authors following DN and DE is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DE wrote the introduction and the section about containerisation and the Rocker Project, and reviewed all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
RC contributed to the section on interfaces for Docker in R (_dynverse_ and `dynwrap`).
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
EH contributed content on \pkg{dockyard}.
LK contributed content on \pkg{dockr}.
SL contributed content on RStudio's usage of Docker.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on the T-Mobile use case.
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
PS wrote the paragraphs about GitLab CI/CD pipelines with Shinyproxy and OpenFaaS templates.
TS \& JW wrote the section on CyVerse.
CvP wrote the section on the usage of Docker containers in Dodona.
CW wrote the sections on Whole Tale and contributed content about publication reproducibility audits.
NX contributed content on \pkg{liftr}.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page} and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide details on the respective contributions.

@@

DN is supported by the project Opening Reproducible Research ([o2r](https://www.uni-muenster.de/forschungaz/project/12343)) funded by the German Research Foundation (DFG) under project number [PE&nbsp;1632/17-1](https://gepris.dfg.de/gepris/projekt/415851837).
The funders had no role in data collection and analysis, decision to publish, or preparation of the manuscript.
KR was supported in part  by a grant from The Leona M. and Harry B. Helmsley Charitable Trust, award number 2016PG-BRI004.
LS and NT are supported by US NIH / NHGRI awards U41HG00405 and U24HG010263.
CW is supported by the Whole Tale project (https://wholetale.org) funded by the US National Science Foundation (NSF) under award [OAC-1541450](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1541450).
NR is supported in part by the Chan-Zuckerberg Initiative Essential Open Source Software for Science program.
We would like to thank Celeste R. Brennecka from the Science Writing Support Service of the University of Münster for her editorial support.

\bibliography{RJreferences}

>>>>

First, different packages use containers for the **deployment of R and \pkg{Shiny} apps**.
\pkg{Shiny} is a popular package for creating interactive online dashboards with R, and it enables users with very diverse backgrounds to create stable and user-friendly web applications \citep{cran_shiny}.
_ShinyProxy_ ([https://www.shinyproxy.io/](https://www.shinyproxy.io/)) is an open-source tool to deploy Shiny apps in an enterprise context, where it features single sign-on, but it can also be used in scientific use cases \citep[e.g., ][]{savini_epiexplorer_2019,glouzon_structurexplor_2017}.
ShinyProxy uses Docker containers to isolate user sessions and to achieve scalability for multi-user scenarios with multiple apps.
ShinyProxy itself is written in Java to accommodate corporate requirements and may itself run in a container for stability and availability.
The tool is built on ContainerProxy ([https://www.containerproxy.io/](https://www.containerproxy.io/)), which provides similar features for executing long-running R jobs or interactive R sessions.
The started containers can run on a regular Docker host but also in clusters.
Continuous integration and deployment (CI/CD) for Shiny applications using Shinyproxy can be achieved, e.g., via GitLab pipelines or with a combination of GitHub and Docker Hub.
A pipeline can include building and checking R packages and Shiny apps.
After the code has passed the checks, Docker images are built and pushed to the container registry.
The pipeline finishes with triggering a [webhook](https://github.com/analythium/shinyproxy-1-click/blob/master/digitalocean/setup.md#setting-up-webhook) on the server, where the
deployment script is executed.
The script can update configurations or pull the new Docker images.
There is a [ShinyProxy 1-Click App](https://marketplace.digitalocean.com/apps/shinyproxy) in the DigitalOcean marketplace that is set up with
these webhooks. The [documentation](https://github.com/analythium/shinyproxy-1-click/blob/master/digitalocean/secure.md) explains how to set up HTTPS with ShinyProxy and webhooks.

@@

The ordering of authors following DN and DE is alphabetical.
DN conceived the article idea, \href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team}, wrote sections not mentioned below, and revised all sections.
DE wrote the introduction and the section about containerisation and the Rocker Project, and reviewed all sections.
DB wrote the section on \pkg{outsider}.
GD contributed the CARD.com use case.
RC contributed to the section on interfaces for Docker in R (_dynverse_ and `dynwrap`).
DC contributed content on Gigantum.
ME contributed to the section on processing and deployment to cloud services.
CF wrote paragraphs about \pkg{r-online}, \pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}.
EH contributed content on \pkg{dockyard}.
LK contributed content on \pkg{dockr}.
SL contributed content on RStudio's usage of Docker.
BM wrote the section on research compendia and made the project Binder-ready.
HN \& JN co-wrote the section on the T-Mobile use case.
KR wrote the section about \pkg{holepunch}.
NR wrote paragraphs about shared work environments and GPUs.
LS \& NT wrote the section on Bioconductor.
PS wrote the paragraphs about CI/CD pipelines with Shinyproxy 1-Click app and OpenFaaS templates.
TS \& JW wrote the section on CyVerse.
CvP wrote the section on the usage of Docker containers in Dodona.
CW wrote the sections on Whole Tale and contributed content about publication reproducibility audits.
NX contributed content on \pkg{liftr}.
All authors approved the final version.
This articles was collaboratively written at \href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The \href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page} and \href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues} provide details on the respective contributions.

@@

DN is supported by the project Opening Reproducible Research ([o2r](https://www.uni-muenster.de/forschungaz/project/12343)) funded by the German Research Foundation (DFG) under project number [PE&nbsp;1632/17-1](https://gepris.dfg.de/gepris/projekt/415851837).
The funders had no role in data collection and analysis, decision to publish, or preparation of the manuscript.
KR was supported in part  by a grant from The Leona M. and Harry B. Helmsley Charitable Trust, award number 2016PG-BRI004.
LS and NT are supported by US NIH / NHGRI awards U41HG00405 and U24HG010263.
CW is supported by the Whole Tale project (https://wholetale.org) funded by the US National Science Foundation (NSF) under award [OAC-1541450](https://www.nsf.gov/awardsearch/showAward?AWD_ID=1541450).
NR is supported in part by the Chan-Zuckerberg Initiative Essential Open Source Software for Science program.
We would like to thank Celeste R. Brennecka from the Scientific Editing Service of the University of Münster for her editorial support.

\bibliography{rockerverse}
